---
title: "iemisc: Additional Examples from GNU Octave size Compatible Functions"
author: "Irucka Embry, E.I.T. (EcoC²S)"
date: "`r Sys.Date()`"
lang: en-us
urlcolor: blue
output:
  rmarkdown::pdf_document:
    highlight: kate
    toc: true
    latex_engine: xelatex
vignette: >
  %\VignetteIndexEntry{iemisc Additional Examples from GNU Octave size Compatible Functions}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

\bigskip

# size Examples (R style)

```{r, warning = FALSE, message = FALSE, tidy = TRUE}

library("iemisc")

import::from(gsubfn, list)
import::from(ramify, mat)

# Example from GNU Octave ndims function reference

size(matlab::ones(4, 1, 2, 1))


# Examples from GNU Octave size function reference

object1 <- matrix(c(1, 2, 3, 4, 5, 6), nrow = 3, ncol = 2, byrow = TRUE)

size(object1)


list[nr, nc] <- size(matrix(c(1, 2, 3, 4, 5, 6), nrow = 3, ncol = 2,
                byrow = TRUE)); nr; nc

size(matrix(c(1, 2, 3, 4, 5, 6), nrow = 3, ncol = 2, byrow = TRUE), 2)


# using ramify's mat

size(mat("1, 2; 3, 4; 5, 6"))

size(mat("1, 2; 3, 4; 5, 6"), 2)

list[nr, nc] <- size(mat("1, 2; 3, 4; 5, 6")); nr; nc

size(matlab::ones(4, 3, 4, 8), 4)

size(matlab::ones(4, 3, 4, 5), 3)

## The following can't be done currently with this function:

# list[nr, remainder] <- size(matlab::ones(2, 3, 4, 5)); nr; remainder

## As a work around to get similar results to GNU Octave, do the following:

nr <- size(matlab::ones(2, 3, 4, 5), 1); nr

remainder <- size(matlab::ones(2, 3, 4, 5), 2); remainder




# Examples from pracma size

size(1:8)

size(matrix(1:8, 2, 4))

size(matrix(1:8, 2, 4), 2)

size(matrix(1:8, 2, 4), 3)

ss <- "object"

size(ss)
```

\bigskip
\bigskip

# size Examples (GNU Octave style)

```{octave}

% check against GNU Octave

% Example from GNU Octave ndims function reference

size(ones(4, 1, 2, 1))


% Examples from GNU Octave size function reference

object1 = [1, 2; 3, 4; 5, 6]

size(object1)


[nr, nc] = size([1, 2; 3, 4; 5, 6])

size([1, 2; 3, 4; 5, 6], 2)

size([1 2; 3 4; 5 6])

size([1 2; 3 4; 5 6], 2)

[nr, nc] = size([1, 2; 3, 4; 5, 6])
nr
nc

size(ones(2, 3, 4, 5))

[nr, remainder] = size(ones(2, 3, 4, 5))
nr
remainder

size(ones(4, 3, 4, 8), 4)

size(ones(4, 3, 4, 5), 3)




% Examples from pracma size

size(1:8)

object2 = [1 3 5 7; 2 4 6 8]

size(object2)

size(object2, 2)

size(object2, 3)

ss = 'object'

size(ss)
```

\bigskip
\bigskip

# length_octave Examples (R style)

```{r, warning = FALSE, message = FALSE, tidy = TRUE}

library("iemisc")

import::from(matlab, ones)

# Example from pracma isempty

object1 <- matrix(0, 1, 0)

length_octave(object1)

object2 <- 2

length_octave(object2)

object3 <- 1:10

length_octave(object3)

object4 <- ones(3, 4)

length_octave(object4)

object5 <- "ss"

length_octave(object5)

object6 <- list(letters, b <- 2)

length_octave(object6)
```

\bigskip
\bigskip

# length Examples (GNU Octave style)

```{octave}

% check against GNU Octave

object1 = [];

length(object1)


object2 = 2;

length(object2)


object3 = 1:10;

length(object3)


object4 = ones(3, 4);

length(object4)


object5 = 'ss';

length(object5)
```

\bigskip
\bigskip

# numel Examples (R style)

```{r, warning = FALSE, message = FALSE, tidy = TRUE}

library("iemisc")

import::from(matlab, ones)

xx <- list(1:26, 1:10)

numel(xx)


# Examples from GNU Octave numel

a <- 1

b <- ones(2, 3)

numel(a, b)


a <- 2

b <- ones(2, 3)

c <- ones(3, 4)

numel(a, b)

numel(a, b, c)

f <- matrix(c(10, 12, 23, 21, 62, 93), nrow = 2, ncol = 3, byrow = TRUE)

g <- c(2, 4)

numel(f, g)
```

\bigskip
\bigskip

# numel Examples (GNU Octave style)

```{octave}

% check against GNU Octave

xx = {1:26, 1:10}


\% Examples from GNU Octave numel
a = 1;

b = ones(2, 3);

numel(a, b)


a = 2;

b = ones(2, 3);

c = ones(3, 4);

numel(a, b)

numel(a, b, c)


f = [10 12 23; 21 62 93];

g = [2 4];

numel(f, g)

```

\bigskip
\bigskip

# ndims Examples (R style)

```{r, warning = FALSE, message = FALSE, tidy = TRUE}

library("iemisc")

# Examples from GNU Octave ndims

b <- matlab::ones(c(4, 1, 2, 1))

ndims(b)

```

\bigskip
\bigskip

# ndims Examples (GNU Octave style)

```{octave}

% check against GNU Octave

% Example from GNU Octave ndims

ndims(ones(4, 1, 2, 1))

```

\bigskip
\bigskip

# isrow Examples (R style)

```{r, warning = FALSE, message = FALSE, tidy = TRUE}

library("iemisc")

# Examples

xx <- ramify::mat("1, 2"); xx

isrow(xx)


xy <- ramify::mat("1, 2; 3, 4"); xy

isrow(xy)
```

\bigskip
\bigskip

# isrow Examples (GNU Octave style)

```{octave}

% check against GNU Octave

isrow([1 2])

isrow([1 2; 3 4])

```

\bigskip
\bigskip

# iscolumn Examples (R style)

```{r, warning = FALSE, message = FALSE, tidy = TRUE}

library("iemisc")

# Examples

xxx <- ramify::mat("1, 2"); xxx

iscolumn(xxx)



xyy <- ramify::mat("1; 2"); xyy

iscolumn(xyy)
```

\bigskip
\bigskip

# iscolumn Examples (GNU Octave style)

```{octave}

% check against GNU Octave

iscolumn([1 2])

iscolumn([1; 2])
```

\bigskip
\bigskip

## Works Cited

Design Guide No. 1110-1-3: Air Stripping Engineering and Design Appendix D: Example Air Stripping By Packed Column, Department Of The Army U.S. Army Corps of Engineers, 31 October 2001, pages D-1 - D-18, http://www.publications.usace.army.mil/Portals/76/Publications/EngineerDesignGuides/DG_1110-1-3.pdf?ver=2013-08-16-101222-003.  

\bigskip
\bigskip

# EcoC²S Links

EcoC²S Home -- https://www.ecoccs.com/  
About EcoC²S -- https://www.ecoccs.com/about_ecoc2s.html  
Products -- https://www.questionuniverse.com/products.html  
EcoC²S Media -- https://www.ecoccs.com/media.html  
EcoC²S Resources -- https://www.ecoccs.com/resources.html  
R Trainings and Resources provided by EcoC²S (Irucka Embry, E.I.T.) -- https://www.ecoccs.com/rtraining.html  

\bigskip
\bigskip

# Copyright and License

All R code written by Irucka Embry is distributed under the GPL-3 (or later) license, see the [GNU General Public License {GPL} page](https://www.gnu.org/licenses/gpl-3.0.html).  

All written content originally created by Irucka Embry is copyrighted under the Creative Commons Attribution-ShareAlike 4.0 International License. All other written content retains the copyright of the original author(s).  


```{r, echo = FALSE, out.width = '100%'}
linguisticsdown::include_graphics2("https://i.creativecommons.org/l/by-sa/4.0/88x31.png")
```

This work is licensed under a [Creative Commons Attribution-ShareAlike 4.0 International License](https://creativecommons.org/licenses/by-sa/4.0/).  
