---
title: "iemisc: Additional Examples from GNU Octave Compatible cosd Function"
author: "Irucka Embry, E.I.T. (EcoC²S)"
date: "`r Sys.Date()`"
lang: en-us
urlcolor: blue
output:
  rmarkdown::pdf_document:
    highlight: kate
    toc: true
    latex_engine: xelatex
vignette: >
  %\VignetteIndexEntry{iemisc Additional Examples from GNU Octave Compatible cosd Function}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

\bigskip

# cosd Examples (R style)

```{r, warning = FALSE, message = FALSE, tidy = TRUE}

library("iemisc")

# Examples from GNU Octave cosd

cosd(c(0, 180, 360))


cosd(c(90, 270, 45))



try(cosd(pi * seq(0, 80, by = 10) / 180))
# gives error message since the computed value is in radians rather than degrees


cos(pi * seq(0, 80, by = 10) / 180)
# this is correct since `cos` expects the angle in radians

try(cosd(seq(0, 80, by = 10) * 180 / pi))
# converts angles in radians to degrees; however, it will still receive an error message with this current implementation. You can use the work-around below:

xx <- seq(0, 80, by = 10) * 180 / pi

cosd(xx)


try(cos(seq(0, 80, by = 10) * 180 / pi))
# converts angles in radians to degrees; however, this is incorrect since `cos` expects the angle in radians and not degrees


cosd(90)

cos(pi/2)
```

\bigskip
\bigskip

# cosd and cos Examples (GNU Octave style)

```{octave}

cosd(90)

cos(pi/2)
```

\bigskip
\bigskip

## Works Cited

Design Guide No. 1110-1-3: Air Stripping Engineering and Design Appendix D: Example Air Stripping By Packed Column, Department Of The Army U.S. Army Corps of Engineers, 31 October 2001, pages D-1 - D-18, http://www.publications.usace.army.mil/Portals/76/Publications/EngineerDesignGuides/DG_1110-1-3.pdf?ver=2013-08-16-101222-003.  

\bigskip
\bigskip

# EcoC²S Links

EcoC²S Home -- https://www.ecoccs.com/  
About EcoC²S -- https://www.ecoccs.com/about_ecoc2s.html  
Products -- https://www.questionuniverse.com/products.html  
EcoC²S Media -- https://www.ecoccs.com/media.html  
EcoC²S Resources -- https://www.ecoccs.com/resources.html  
R Trainings and Resources provided by EcoC²S (Irucka Embry, E.I.T.) -- https://www.ecoccs.com/rtraining.html  

\bigskip
\bigskip

# Copyright and License

All R code written by Irucka Embry is distributed under the GPL-3 (or later) license, see the [GNU General Public License {GPL} page](https://www.gnu.org/licenses/gpl-3.0.html).  

All written content originally created by Irucka Embry is copyrighted under the Creative Commons Attribution-ShareAlike 4.0 International License. All other written content retains the copyright of the original author(s).  


```{r, echo = FALSE, out.width = '100%'}
linguisticsdown::include_graphics2("https://i.creativecommons.org/l/by-sa/4.0/88x31.png")
```

This work is licensed under a [Creative Commons Attribution-ShareAlike 4.0 International License](http://creativecommons.org/licenses/by-sa/4.0/).  
