---
title: "iemisc: Comparing Other Hydraulic Software Output to iemisc's Manningtrap for Critical Conditions"
author: "Irucka Embry, E.I.T. (EcoC²S)"
date: "`r Sys.Date()`"
lang: en-us
urlcolor: blue
output:
  rmarkdown::pdf_document:
    highlight: kate
    toc: true
    latex_engine: xelatex
vignette: >
  %\VignetteIndexEntry{iemisc Comparing Other Hydraulic Software Output to iemisc's Manningtrap for Critical Conditions}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

\bigskip

# Replicate the R code without the images

Note: If you wish to replicate the R code below, then you will need to copy and paste the following commands in R first (to make sure you have all the of the required packages):  

```{r eval = FALSE, tidy = TRUE}
install.packages("iemisc", "knitr")
# install the package and its dependencies
```

\bigskip
\bigskip

```{r, warning = FALSE, message = FALSE, tidy = TRUE}
# load the required package
library("iemisc")
```

\bigskip
\bigskip

# FHWA Hydraulic Toolbox Example 1

```{r, echo = FALSE, out.width = '100%'}
linguisticsdown::include_graphics2("compare.png")
```

\bigskip
\bigskip

# Dr. Xing Fang's Open Channel Flow Calculator's Solution of Example 1

```{r, echo = FALSE, out.width = '100%'}
linguisticsdown::include_graphics2("trapezoid1.png")
```

\bigskip
\bigskip

# iemisc's Manningtrap Solution of Example 1

```{r, warning = FALSE, message = FALSE, tidy = TRUE}

uuc1 <- Manningtrap_critical(y = 1.454, b = 4, m = 3, Sf = 0.02, n = 0.0550, units = "Eng", type = "symmetrical", critical = "accurate", output = "data.table")

knitr::kable(uuc1)
```

\bigskip
\bigskip
\bigskip
\bigskip

# FHWA Hydraulic Toolbox Example 2

```{r, echo = FALSE, out.width = '100%'}
linguisticsdown::include_graphics2("2019-05-22_08_24_13-ChannelAnalysis.png")
```

\bigskip
\bigskip

# Dr. Xing Fang's Open Channel Flow Calculator's Solution of Example 1

```{r, echo = FALSE, out.width = '100%'}
linguisticsdown::include_graphics2("trapezoid2.png")
```

\bigskip
\bigskip

# iemisc's Manningtrap Solution of Example 2

```{r, warning = FALSE, message = FALSE, tidy = TRUE}

uuc2 <- Manningtrap_critical(Q = 745, b = 8, m1 = 2, m2 = 1.5, Sf = 0.01, n = 0.0150, units = "Eng", type = "non-symmetrical", critical = "accurate", output = "data.table")

knitr::kable(uuc2)
```

\bigskip
\bigskip
\bigskip
\bigskip

# Works Cited

FHWA Hydraulic Toolbox Version 4.4. https://www.fhwa.dot.gov/engineering/hydraulics/software/toolbox404.cfm  

The open channel flow calculator. Dr. Xing Fang, Department of Civil Engineering, Lamar University, 2000. https://eng.auburn.edu/~xzf0001/Handbook/Channels.html  

\bigskip
\bigskip

# EcoC²S Links

EcoC²S Home -- https://www.ecoccs.com/  
About EcoC²S -- https://www.ecoccs.com/about_ecoc2s.html  
Products -- https://www.questionuniverse.com/products.html  
EcoC²S Media -- https://www.ecoccs.com/media.html  
EcoC²S Resources -- https://www.ecoccs.com/resources.html  
R Trainings and Resources provided by EcoC²S (Irucka Embry, E.I.T.) -- https://www.ecoccs.com/rtraining.html  

\bigskip
\bigskip

# Copyright and License

All R code written by Irucka Embry is distributed under the GPL-3 (or later) license, see the [GNU General Public License {GPL} page](https://www.gnu.org/licenses/gpl-3.0.html).  

All written content originally created by Irucka Embry is copyrighted under the Creative Commons Attribution-ShareAlike 4.0 International License. All other written content retains the copyright of the original author(s).  


```{r, echo = FALSE, out.width = '100%'}
linguisticsdown::include_graphics2("https://i.creativecommons.org/l/by-sa/4.0/88x31.png")
```

This work is licensed under a [Creative Commons Attribution-ShareAlike 4.0 International License](http://creativecommons.org/licenses/by-sa/4.0/).  
